\Large\bf Simultaneous Elicitation of Committee and Voters Preferences

### 2019-03-29 RJCIA2019/short_draft.tex
RJCIA2019

\input{ijcai_preamble} 
\author{ID: 2696}
%\author{Beatrice Napolitano$^1$\and Olivier Cailloux$^1$	\and  Paolo Viappiani$^{2}$
%	\affiliations $^1$ LAMSADE, UMR 7243, CNRS and Universit\'e Paris Dauphine, PSL Research University, Paris, France\\ $^2$ LIP6, UMR 7606, CNRS and Sorbonne Universit\'e, Paris, France
%	\emails \{firstname.lastname\}@dauphine.fr, paolo.viappiani@lip6.fr
%}
%Restoring page numbers as missing them out violates basic human rights of the reviewers. -- Olivier
\thispagestyle{plain}
\pagestyle{plain}

@@

@@

\title{Robust Winner Determination and Simultaneous Elicitation of Scoring Rules and Preferences}

\begin{document}
\maketitle
\begin{abstract}
Social choice deals with the problem of determining a consensus choice from the preferences of different agents (voters).
In the classical setting, the voting rule is fixed beforehand and full information concerning the preferences of the agents is provided.
Recently, the assumption of full preference information has been questioned by a number of researchers and several methods for eliciting preferences have been proposed.

In this paper we go one step further and we assume that both the voting rule and the agents preferences are partially specified.
Focusing on positional scoring rules, we assume that the chair, while  not able to give a precise definition of the rule, is capable of answering simple questions  requiring to pick a winner from a specific  example profile.
Moreover, the preferences of the agents are incrementally acquired by asking comparison queries.

We propose a method  for robust approximate winner determination in this setting with minimax regret. 
We then provide an interactive elicitation protocol based on minimax regret
and develop several query strategies that interleave questions to the chair and questions to the agents in order to acquire the most relevant information in order to quickly converge to optimal or a near-optimal alternative.
\end{abstract}

\section{Introduction}
Aggregation of preference information is a central task in many computer systems (recommender systems, search engines, etc).
In many situations, such as in group recommender systems, the goal is to find a consensus choice.
It is therefore natural to look at methods from social choice and see how they can be adapted for group decision-making in a computerized setting.

The traditional approach to social choice assumes that both the social function and the full preference orderings of the agents are expressed beforehand. This represents two strong hypothesis.
About the second one, requiring agents to express full preference orderings can be prohibitively costly (in term of cognitive and communication cost), especially for decisions with large sets of alternatives.
This observation has motivated a number of recent works considering social choice with partial preference orders \citep{Xia2008, Pini2009, Konczak05} and incremental elicitation \citep{Kalech2011, Lu2011, Naamani-Dery2015} of agents’ preferences.

The first hypothesis also should be relaxed, for in several situations it may not be easy to precisely define the voting rule.
Indeed, it is possible that the chair may have some preferences over the desired aggregation method, but may still be not able to formalize the voting rule upfront.
The work of \citet{Cailloux2014} provides elicitation methods for a quite general class of rules based on weak orders.
When considering positional scoring rules, several authors \citep{Stein1994, Llamazares2013, Viappiani2018} have worked on positional scoring rules with uncertain weights, assuming that the preferences of the agents are fully known.
%it is possible to derive dominance relations (akin to stochastic dominance) that allow to eliminate some alternatives since they will be less preferred than another one for any instantiation of the weights \citep{Stein1994}.
%Among others, Llamazares and Pe{\~{n}}a have considered  the problem of dealing with underspecified weights in positional scoring rules.

In this paper we consider that both the agents’ preferences and the social choice rule are partially specified.
We develop methods for computing the minimax-optimal alternative for scoring rules with partial agent preference information and partial information about the weights of the scoring rule.
While previous works have considered either partial information about the agents preferences or a partially specified aggregation method, we do not know of any work considering both sources of uncertainty at the same time.

In this work, we focus on positional scoring rules, that  are a particularly common method used to aggregate rankings and determine a winner.
We also address the problem of elicitation, providing incremental elicitation methods to acquire relevant preference information; we discuss  several heuristics that determine queries that quickly reduce minimax regret.
In particular, our query strategies focus simultaneously on reduction of relevant preference and voting rule uncertainty.

The paper is organized as follows.
In Section \ref{sec:background} we provide the necessary background.
We introduce the minimax criterion, that selects a winner that minimizes the worst possible regret, in Section \ref{sec:mmr}.
In Section \ref{sec:elicit} we provide an interactive elicitation protocol based on minimal regret;  in Section \ref{sec:experiments} we present the empirical validation of our approach with simulations; and Section \ref{sec:conclusions} provides some final thoughts.

\section{Social choice with partial information}
\label{sec:background}
We now introduce some basic concepts.
We assume defined a set $A$ of $m$ alternatives (products, restaurants, movies, public projects, job candidates, etc.) and a set $\set{1, …, n}$ of agents (voters); each agent $j$ is associated to her “real” preference order $\pref_j$. The agents come from an infinite set $\N$ of potential agents.
The real preferences of the agents are elements of $\linors$, the linear orders (connected, transitive, asymmetric relations) over the alternatives.
Following the social choice nomenclature, we call {\em profile} the association of a preference to each agent, considering a subset of agents from the set $\N$, and denote a profile by $(\pref_1,\ldots,\pref_n)$.
A profile is equivalently represented by $\profile=(v_1,\ldots,v_n)$ where $v_j(i) \in \set{1, \ldots, m}$ denotes the rank (position) of alternative $i$ in the preference order $\pref_j$.

Let $V$ be the set of possible preference profiles (the union for any integer $n$ of the $n$-fold cartesian product of the linear orders over the alternatives).
A social choice function $f : V \rightarrow \powersetz{A}$ associates a profile with a set of winners, where $\powersetz{A}$ represents the set of subsets of $A$ except for the emptyset. (Sets are used for tied winners.)
Among the many possible social choice functions, we consider {\em positional scoring rules (PSR)}, which attach weights to positions according to the vector $(w_1, \ldots, w_m)$ (also called the scoring vector).
An alternative obtains a score that depends on the rank obtained in each of the preference orders:
\begin{align}
\label{eq:srule}
s(x; \profile, \w) = \sum_{j=1}^{n} w_{v_j(x)}
= \sum_{r=1}^{m} \alpha^{x}_r w_r 
\end{align}
where $\alpha^{x}_r$ is the number of times that alternative $x$ was ranked in the $r$-th position.
The winners are the alternatives with highest scores.

In this work we assume fixed, but unknown, a profile $\profile^*$ (and corresponding agents preferences $\pref_j^*$) and a weight vector $\w^*$, and we want to reason about partial preference information concerning those objects.
Our knowledge at a given time of the preference of agent $j$ is encoded by a partial order over the alternatives, thus a transitive and asymmetric binary relation, denoted by $\ppref_j$. 
In this work we assume that preference information is truthful, i.e. $a \ppref_j b ⇒ a \pref_j^* b$.
We use $\pinc$ to denote incomparability, that is $a \pinc_{j} b$ iff $a \nppref_j b \wedge b \nppref_j a$.
An incomplete profile $\pprofile = (\ppref_1, \ldots, \ppref_n)$ maps each agent to a partial preference.

A completion of $\ppref_j$ is any linear order $\pref$ that extends $\ppref_j$.
Let $C(\ppref_j) = \set{{\succ} \in \linors \suchthat {\ppref_j} \subseteq {\succ}}$ be the set of possible completions of $\ppref_i$.
We let $C(\pprofile)=C(\ppref_1)\times … \times C(\ppref_n)$ be the set of complete profiles extending $p$. Note that $\profile^* \in C(\pprofile)$.

\medskip
We also assume that the weights of the scoring rule are only partially specified.
Therefore the vector $(w_1,\ldots,w_m)$  is not known but we are given a set of constraints restraining the possible values that weights can take.
We consider a decreasing sequence of weights:
\begin{align}
1=w_{1} > w_{2} > \ldots > w_{m}=0. \label{eq:monotone}
\end{align}
This is a natural assumption, as it is better to be ranked first than second, second than third, etc. 
Without loss of generality, we consider that $w_1=1$ and $w_m=0$.

The weights of a scoring rule can model different preferences of the chair. 
For instance, the weights can control the inclination to favor ``extreme'' alternatives (often at either the top or the bottom of the input rankings) at the expenses of ``moderate'' alternatives (that are more consistently in the middle part of the input rankings).

An important class of scoring rule is the one composed of weights that are a convex sequence \citep{Stein1994,Llamazares2016}, meaning that the difference between the weight of the first position and the weight of the second position is at least as great as the difference between the weights of the second and third positions, etc.
\begin{equation} 
\label{eq:convexity}
\forall r \in \{1,\ldots,m-2\}: w_r - w_{r+1} \geq w_{r+1}-w_{r+2}.
\end{equation}
The constraint above is often used when aggregating rankings in sport competitions.
We use $\W$ to denote the set of convex weight vectors.

In general it can be difficult to set the weights in an appropriate way.
We assume that in addition of basic requirements (monotonicity and convexity), the chair (the person or the organization that is supervising the voting process) is able to specify additional preferences about how the social choice function should behave.
In this work we assume that preferences of the chair are encoded with linear constraints about the vector $\w$, relating the value of the weights of different positions.
We use $\pw \subseteq \W$ to denote the set of weight vectors compatible with the preferences expressed by the chair about the scoring vector, and $\Co_W$ to denote the set of linear constraints expressed by the chair.

Of course it may be difficult for real decision makers to state preferences about the voting rule in such an abstract way.
These additional preferences can be elicited by asking questions about concrete profiles, for instance, by showing a complete profile of a small synthetic election and asking who should be elected in this case, as shown in Section \ref{sec:elicit}.

\section[Minimax regret under partial profile and weight information]{
Robust winner determination}
\label{sec:mmr}
In this paper, we consider a setting where both the agents' preferences and the preferences of the chair about the voting rule are incomplete.
Notice that some authors have considered possible and necessary winners assuming a partial profile  \citep{Xia2008} or assuming an incompletely specified scoring rule \citep{Viappiani2018};
however, we note that, in typical settings, there will be no necessary winner and too many possible winners.
In practice, however, it is often useful to declare a winner given partial information.

As a decision criterion to determine a winner, we propose to use minimax regret. 
Minimax regret \citep{Savage1954} is a decision criterion that has been used for robust optimization under data uncertainty \citep{Kouvelis1997} as well as in decision-making with uncertain utility values \citep{Salo2001,Boutilier2006}.
\citet{Lu2011} have adopted minimax regret for winner determination in social choice with
the preferences of the agents that are only partially known, while the social choice function is predetermined and known.

In this work we consider the simultaneous presence of uncertainty in the agents' preferences and uncertainty in the weights.
Using {\em maximum regret} to quantify the worst-case error, the alternative that minimize this error is selected as winner, providing us with a form of robust optimization.
Intuitively, the quality of a proposed alternative $a$ is how far from optimal $a$  could be in the worst case, given the current knowledge about the voting rule and about the agents' preferences.

The maximum regret is considered by assuming that an adversary both 1) can extend the partial profile $\pprofile$ into a complete profile, and 2) can instantiate the weights choosing among any weight vector in $\pw$, where $\pprofile$ and $\pw$ represent our knowledge so far.
We formalize the notion of minimax regret in multiple steps.
First of all, $\Regret(x, \profile, \w)$ is the “regret” of selecting $x$ as a winner instead of choosing the optimal alternative under $\profile$ and $\w$:
\[\Regret(x, \profile, \w) = \max_{y \in A} s(y; \profile,\w) - s(x; \profile, \w).\]
The pairwise max regret $\PMR(x,y;\pprofile,W)$ of $x$ relative to $y$ given partial profile $\pprofile$ and the set of weights $W$
is the worst-case loss of choosing $x$ instead of $y$ under all possible realizations of the full profile {\em and} all possible instantiations of the weights:
\begin{align}
\PMR(x,y; \pprofile, W) & = \max_{\w \in W} \max_{\profile \in C(\pprofile)} s(y; \profile,\w) - s(x; \profile,\w).
\end{align}

Max regret $\MR(x;\pprofile,W)$ is the worst-case loss of $x$. It is the loss occurred by an adversarial selection of a complete profile $\profile$ extending $\pprofile$ and a selection of $\w \in W$ to maximize the loss between $x$ and the true winner under $\profile$ and $\w$.
\begin{align}
\MR(x; \pprofile, W) & = \max_{y \in A} \PMR(x,y; \pprofile, W)\\
& = \max_{\w \in W} \max_{\profile \in C(\profile)} \Regret(x, \profile, \w).
\end{align}

Finally,  $\MMR(\pprofile,W)$ is the value of minimax regret under $\pprofile$ and $W$, obtained when recommending a minimax optimal alternative $x^*_{\pprofile, W}$:
\begin{align*}
\MMR(\pprofile,W) & = \min_{x \in A} \MR(x;\pprofile,W) \\
x^{*}_{\pprofile,W} \in A^*_{\pprofile, W} & = \argmin_{x \in A} \MR(x;\pprofile,W) 
\end{align*}
By picking as consensus choice
 an alternative associated with minimax regret, we can provide a recommendation that gives worst-case guarantees, giving some robustness in face of uncertainty (due to both not knowing the agents' preferences and the weights used in the aggregation). 
In cases of ties in minimax regret, we can either decide to return all minimax alternatives $A^*_{\pprofile, W}$ as winners or to pick just one of them using some tie-breaking strategy.

Notice that if $\MMR(\pprofile, W)\!=\!0$, then $x^{*}_{\pprofile,W}$ is a necessary co-winner; this means that for any valid completion of the profile and any feasible $w \!\in\! W$, $x^{*}_{\pprofile,W}$ obtains a highest score.

We note that our notion of regret gives some cardinal meaning to the scores: instead of just being used to select winners under the corresponding PSR, their differences are considered as representing the regret of the chair.

% Add some general remarks about using minimax regret

\paragraph{Computation of minimax regret}
In order to compute pairwise maximum regret and therefore minimax regret, we adapt the reasoning from \citet{Lu2011} so that we decompose the $\PMR$ into the contributions associated to each agent.
The setting is however more challenging due to the presence of uncertainty in the weights.

Scoring rules are additively decomposable.
%A scoring rule associates each alternative to a score that is given by the sum of points obtained for each agent.
Recall that when computing $s(x; \profile, \w)$, $w_{v_j(x)}$ is the number of points that $x$ obtains in the ranking $v_j$ (see Equation \ref{eq:srule}).
By exploiting the decomposition of the score in terms of votes, we write the actual regret of choosing $x$ instead of $y$ as
\[
s(y; \profile,\w) - s(x; \profile, \w) = \sum_{j=1}^n w_{v_j(y)} - w_{v_j(x)},
\]
and  can rewrite $\PMR$ as follows:
\begin{align*}
& \PMR(x,y; \pprofile, W) = \max_{\w \in W} \max_{\profile \in C(\pprofile)} [ s(y; \profile,\w) - s(x; \profile,\w) ] \\
& =  \max_{\w \in W} \sum_{j=1}^{n} \max_{v_j \in C(\succ_j^p)} [w_{v_j(y)} - w_{v_j(x)}]. \\
\end{align*}
Note that in general the inner max will depend on the weights chosen by the outer min.

We consider now a procedure for completing a partial profile that was first proposed by \citet{Lu2011} when considering minimax regret with a fixed weight vector.
As we will show, this procedure can also be used in our case of uncertain weight vector.

We are interested in computing $\PMR(x, y; \pprofile, W)$. This represents the “worst” difference of score, thus the difference of score between $y$ and $x$ under the worst case preferences compatible with $\pprofile$ and $W$, where the worst case is the one that maximizes this difference of score.
Consider our knowledge $\ppref_j$ of the preference of an agent $j$. 
The adversary has to make the score of $y$ as high as possible, and the score of $x$ as low as possible. 
To do this, he should complete $\ppref_j$ to $\pref_j$ by putting above $x$ as many alternatives as he can, that is, all the alternatives except those that are known to be worse than $x$ (those $a$ such that $x \ppref_j a$); and similarly, put below $y$ all the alternatives he can, thus, put above $y$ only those $a$ such that $a \ppref_j y$. 
Sometimes these two objectives conflict: when an alternative should go above $x$ according to the first objective and below $y$ according to the second one (because $¬(x \ppref_j a) ∧ ¬(a \ppref_j y)$), and $x$ is known to be better than $y$ (thus $x \ppref_j y$). 
In this case, the first objective should take priority, and $a$ should be put above $x$, which will move both $x$ and $y$ one rank lower than if $a$ had been put below $y$. 
This maximizes the adversary’s interests: because the weight vector is convex, the difference of scores will be lower when both alternatives are ranked lower (Equation \ref{eq:convexity}), and that difference of scores is in favor of $x$ when $x \ppref_j y$, thus to be minimized from the point of view of the adversary.
To summarize:
\begin{align} 
\label{eq:compl}
\begin{split}
a \pref_j x &⇔ ¬(x \ppref_j a)\\
y \pref_j a &⇔ ¬(a \ppref_j y) ∧ (¬(x \ppref_j y) ∨ (x \ppref_j a)).
\end{split}
\end{align} 
We just sketched a proof of the following claim, and leave the rest of the proof to the reader.
\begin{claim}
There exists a completion $\hat{\profile} \in C(\pprofile)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} [ s(y; \hat{\profile}, \w) - s(x; \hat{\profile}, \w) ]$ and such that the linear order $\hat{v}_{j}$ of each agent $j$ satisfies (\ref{eq:compl}).
\end{claim}
It follows that $\hat{v}_{j}(x)$, the rank of $x$ in the PMR-maximizing linear order of agent $j$, 
is $\hat{v}_{j}(x) = 1+\card{A}-\card{{\ppref_j}(x)}$, where ${\ppref_j}(x)$ designates the set of alternatives that are less good than $x$ according to $\ppref_j$, and the rank of $y$ is $\hat{v}_{j}(y)=1+\card{{\pprefinv_j}(y)}+\beta$, where $\beta = \card{A} - \card{{\ppref_j}(x)}$ if $(x \ppref_j y)$ and $\beta = 0$ otherwise.

%Proof: Define $\pref$ as just $\ppref_j$ with the arcs added to minimally satisfy the above. Thus, add the following arcs: when $x ? a$, add $a > x$. When $a ? y ∧ x \ppref_j y ∧ ¬(x \ppref_j a)$, add $a \pref y$. When $a ? y ∧ [¬(x \ppref_j y) ∨ (x \ppref_j a)]$, add $y \pref a$. Observe that $\ppref_j \subseteq \pref$. We have to show that it is acyclic. (Then we take its transitive closure to make it a partial order, and it’s done.) Note that if there is a cycle in $\pref$, it’s because of some new arcs, as $\ppref_j$ is acyclic. Thus, the cycle must go through one of these new arcs. …

\begin{claim}
The $\PMR$ can be written as:
\begin{align} 
\PMR(x,y; \pprofile, W)  
& = \max_{w \in W} \sum_{j=1}^n w_{\hat{v}_j(y)} - w_{\hat{v}_j(x)} = \\ 
& = \max_{w \in W} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_i. 
\end{align}
where the $\hat{\alpha}_{r}^{y}$ (resp. $\hat{\alpha}_{r}^{x}$)  is the number of times $y$ (resp. $x$) is at rank $r$ in the rankings of the completed profile $\hat{\profile}$.
\end{claim}
The last claim shows that PMR is linear in the weights.
Recall that the preferences of the chair are encoded with linear constraints $\Co_{W}$.
The pairwise max regret $\PMR(x,y; \pprofile,W)$ can be obtained as the solution of the following linear program defined on the variables $w_1, …, w_m$, the weights attached to the different positions. 
\begin{align*}
\max_{\w} & \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_{r}\\
\text{ s.t. } & \text{Equation } (\ref{eq:monotone})\\
& \text{Equation } (\ref{eq:convexity})\\
& \Co_W
\end{align*}
Note that given our choice $w_{1}=1$ and $w_{m}=0$, there are only $m-2$ variables 
(we leave $w_{1}$ and $w_{m}$ in the LP just for clarity of presentation).

The max regret $\MR(x; \pprofile, W)$ is determined by considering the pairwise regret of $x$ with each other alternative in $A$.
Optimal alternatives w.r.t. minimax regret are the ones with least max regret. 
Notice that, whenever the PMR of an alternative $x$ (against some other alternative) exceeds the best MR value found so far, we don't need to further evaluate $x$, as it can't be optimal. 
This idea can be exploited further by adopting a minimax-search tree \citep{Braziunas2011}.

\section{Interactive Elicitation} 
\label{sec:elicit}
We propose an incremental elicitation method based on minimax regret.
At each step, the system may either ask a question to one of the agents, or ask a question to the chair about the voting rule. 
The goal is to acquire relevant information to reduce minimax regret as quickly as possible.
As termination condition of elicitation, we can check whether minimax regret is lower than a threshold; if we wish optimality, we can perform elicitation until minimax regret drops to zero.

The remainder of this Section is structured as follows.
First of all, we discuss the different types of questions that can be asked to the agents and to the chair, and the way responses are handled.
Then, we describe different strategies to determine informative queries to ask next, with the goal of reducing $\MMR(\pprofile,W)$ quickly.

>>>>

%\input{rfia} 
\documentclass[a4paper,twoside]{article}
\usepackage{rfia2000}
\usepackage[T1]{fontenc}
\usepackage{babel}
\usepackage{times}

%\author{ID: 2696}
\author{\begin{tabular}[t]{c@{\extracolsep{6em}}c@{\extracolsep{6em}}c}
	B. Napolitano${}^1$ & O. Cailloux${}^1$ & P. Viappiani${}^2$ \\
	\end{tabular}
{} \\
\\
${}^1$        LAMSADE, UMR 7243, CNRS and Universit\'e Paris Dauphine, PSL Research University, Paris, France   \\
${}^2$        LIP6, UMR 7606, CNRS and Sorbonne Universit\'e, Paris, France
{} \\
\\
%Mon adresse compl\`ete \\
\{firstname.lastname\}@dauphine.fr, paolo.viappiani@lip6.fr\\
%{\bf Domaine principal de recherche}: RFP ou IA\\
%{\bf Papier soumis dans le cadre de la journée commune}: OUI ou NON
}

\thispagestyle{plain}
\pagestyle{plain}

@@

\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{booktabs}

@@

\title{\Large\bf Simultaneous Elicitation of Committee and Voters Preferences}

\begin{document}
\date{}
\maketitle
\thispagestyle{empty}
%\subsection*{R\'esum\'e}
%{\em
%	Le choix social traite du probl\`eme de la d\'etermination d'un choix consensuel \`a partir des pr\'ef\'erences de diff\'erents agents. Dans le cadre classique, la r\`egle de vote est fix\'ee \`a l’avance et les informations compl\`etes concernant les pr\'ef\'erences des agents sont fournies. Dans cet article, nous supposons que la r\`egle de vote et les pr\'ef\'erences des agents sont partiellement connues. Ainsi, nous fournissons un protocole interactif d’\'elicitation bas\'e sur le regret minimax et d\'eveloppons plusieurs strat\'egies d’interrogation, associant questions au comit\'e et questions aux agents, afin d’obtenir les informations les plus pertinentes et de converger rapidement vers un r\'esultat optimal ou s'approchant de l'optimal.
%}
%\subsection*{Mots Clef}
%Choix social computationnel, IA dans l'incertain, \'elicitation des pr\'ef\'erences, regret minimax.

\subsection*{Abstract}
{\em
	Social choice deals with the problem of determining a consensus choice from the preferences of different agents (voters). In the classical setting, the voting rule is fixed beforehand and full information concerning the preferences of the agents is provided. In this paper we go one step further and we assume that both the voting rule and the agents preferences are partially specified. In this setting, we provide an interactive elicitation protocol based on minimax regret and develop several query strategies that interleave questions to the chair and questions to the agents in order to acquire the most relevant information in order to quickly converge to optimal or a near-optimal alternative.
}
\subsection*{Keywords}
Computational Social Choice, uncertainty in AI, preference elicitation, minimax regret.

\section{Introduction}
%Aggregation of preference information is a central task in many computer systems (recommender systems, search engines, etc) whose final goal is to find a consensus choice.
%In this contexts, a natural approach consists in looking at methods from social choice and see how they can be adapted for group decision-making in a computerized setting.
In a traditional social choice setting both the social function and the full preference orderings of the agents are expressed beforehand. Nevertheless, sometimes requiring agents to express full preference orderings can be prohibitively costly (in term of cognitive and communication cost), especially for decisions with large sets of alternatives.
This observation has motivated a number of recent works considering social choice with partial preference orders \citep{Xia2008, Pini2009, Konczak05} and incremental elicitation \citep{Kalech2011, Lu2011, Naamani-Dery2015} of agents’ preferences. 
Furthermore, several situations for which it may not be easy to precisely define the voting rule exist.
Indeed, it is possible that the chair has some preferences over the desired aggregation method, but it is still not able to formalize the voting rule upfront.
The work of Cailloux and Endriss \citep{Cailloux2014} provides elicitation methods for a quite general class of rules based on weak orders.
When considering positional scoring rules, several authors \citep{Stein1994, Llamazares2013, Viappiani2018} have worked on positional scoring rules with uncertain weights, assuming that the preferences of the agents are fully known.
In this paper we consider that both the agents’ preferences and the social choice rule are partially specified.
We develop methods for computing the minimax-optimal alternative using positional scoring rules and we provide incremental elicitation methods to acquire relevant preference information. We then discuss several heuristics that determine queries, either to a voter or to the committee, that quickly reduce minimax regret.
While previous works have considered either partial information about the agents preferences or a partially specified aggregation method, we do not know of any work considering both sources of uncertainty at the same time.

\section{Partial information}
\label{sec:background}
We assume defined a set $A$ of $m$ alternatives (products, restaurants, public projects, etc.) and a set $\set{1, …, n}$ of agents (voters). Each agent $j$ comes from an infinite set $\N$ of potential agents and is associated to her “real” preference order $\pref_j \in \linors$ which is a linear order (connected, transitive, asymmetric relation) over the alternatives.
A {\em profile} is the association of a preference to each agent and it is equivalently represented by $(\pref_1,\ldots,\pref_n)$ or by $\profile=(v_1,\ldots,v_n) \in V$ where $v_j(i) \in \set{1, \ldots, m}$ denotes the rank (position) of alternative $i$ in the preference order $\pref_j$.
%Given $V$ the set of possible preference profiles, a social choice function $f : V \rightarrow \powersetz{A}$ associates a profile with a set of winners; where $\powersetz{A}$ represents the set of subsets of $A$ except for the emptyset.
A social choice function associates a profile with a set of winners; we consider {\em positional scoring rules (PSR)}, which attach weights to positions according to the scoring vector $(w_1, \ldots, w_m)$.
An alternative obtains a score that depends on the rank obtained in each of the preference orders and the winners are the alternatives with highest scores.

In this work we assume fixed, but unknown, a profile $\profile^*$ and a weight vector $\w^*$.
Our knowledge at a given time of the preference of agent $j$ is encoded by a partial order over the alternatives, thus a transitive and asymmetric binary relation, denoted by $\ppref_j$; we assume that preference information is truthful, i.e. $a \ppref_j b ⇒ a \pref_j^* b$.
%We use $\pinc$ to denote incomparability, that is $a \pinc_{j} b$ iff $a \nppref_j b \wedge b \nppref_j a$.
An incomplete profile $\pprofile = (\ppref_1, \ldots, \ppref_n)$ maps each agent to a partial preference. A completion of $\ppref_j$ is any linear order $\pref$ that extends $\ppref_j$ and we indicate with $C(\ppref_j) = \set{{\succ} \in \linors \suchthat {\ppref_j} \subseteq {\succ}}$ the set of possible completions of $\ppref_i$.
%We let $C(\pprofile)=C(\ppref_1)\times … \times C(\ppref_n)$ be the set of complete profiles extending $p$. Note that $\profile^* \in C(\pprofile)$.

Our knowledge regarding the weights of the scoring rule is, instead, represented by
%We also assume that the weights of the scoring rule are only partially specified.
%Therefore the vector $(w_1,\ldots,w_m)$  is not known but we are given 
a set of constraints restraining the possible values that they can take. We consider a convex sequence of weights and we assume that 
%, i.e. a decreasing sequence $1=w_{1} > w_{2} > \ldots > w_{m}=0$ such that the difference between the weights of the first and second positions is at least as great as the difference between the weights of the second and third positions, etc.
%\begin{equation} 
%\label{eq:convexity}
%\forall r \in \{1,\ldots,m-2\}: w_r - w_{r+1} \geq w_{r+1}-w_{r+2}.
%\end{equation}
%We assume that in addition of basic requirements (monotonicity and convexity), 
the chair is able to specify additional preferences about how the social choice function should behave. Of course it may be difficult for real decision makers to state these preferences, thus we can elicit them, for instance, by showing a complete profile of a small synthetic election and asking who should be elected in this case.
Such requirements are encoded with linear constraints about the vector $\w$ and the set of these constraints is denoted by $\Co_W$. We use $\pw \subseteq \W$ to denote the set of weight vectors compatible with the preferences expressed by the chair.

\section[Minimax regret under partial profile and weight information]{
Robust winner determination}
\label{sec:mmr}
As a decision criterion to determine a winner, we propose to use minimax regret \citep{Savage1954} . 
%Minimax regret \citep{Savage1954} is a decision criterion that has been used for robust optimization under data uncertainty \citep{Kouvelis1997} as well as in decision-making with uncertain utility values \citep{Salo2001,Boutilier2006}.
%\citet{Lu2011} have adopted minimax regret for winner determination in social choice with
%the preferences of the agents that are only partially known, while the social choice function is predetermined and known.
%In this work we consider the simultaneous presence of uncertainty in the agents' preferences and uncertainty in the weights.
We use {\em maximum regret} to quantify the worst-case error, which measures, intuitively, how far an alternative is from the optimal one given current knowledge.
% about the voting rule and the agents' preferences.
The alternative that minimize this error is selected as winner, providing us with a form of robust optimization.
The maximum regret is considered by assuming that an adversary can both 1) extend the partial profile $\pprofile$ into a complete one, and 2) instantiate the weights choosing among any weight vector in $\pw$.
We formalize the notion of minimax regret in multiple steps.
First of all, $\Regret(x, \profile, \w)$ is the “regret” of selecting $x$ as a winner instead of choosing the optimal alternative under $\profile$ and $\w$:
\[\Regret(x, \profile, \w) = \max_{y \in A} s(y; \profile,\w) - s(x; \profile, \w).\]
The pairwise max regret $\PMR(x,y;\pprofile,W)$ of $x$ relative to $y$ given partial profile $\pprofile$ and the set of weights $W$ is the worst-case loss of choosing $x$ instead of $y$ under all possible realizations of the full profile {\em and} all possible instantiations of the weights. Max regret $\MR(x;\pprofile,W)$ is the worst-case loss of $x$ and $\MMR(\pprofile,W)$ is the value of minimax regret obtained when recommending a minimax optimal alternative $x^*_{\pprofile, W}$. 
%\begin{align}
%\PMR(x,y; \pprofile, W) & = \max_{\w \in W} \max_{\profile \in C(\pprofile)} s(y; \profile,\w) - s(x; \profile,\w).
%\end{align}
%It is the loss occurred by an adversarial selection of a complete profile $\profile$ extending $\pprofile$ and a selection of $\w \in W$ to maximize the loss between $x$ and the true winner under $\profile$ and $\w$.
\begin{align}
\PMR(x,y; \pprofile, W) & = \max_{\w \in W} \max_{\profile \in C(\pprofile)} s(y; \profile,\w) - s(x; \profile,\w) \\
\MR(x; \pprofile, W) & = \max_{y \in A} \PMR(x,y; \pprofile, W)\\
%& = \max_{\w \in W} \max_{\profile \in C(\profile)} \Regret(x, \profile, \w) \\
\MMR(\pprofile,W) & = \min_{x \in A} \MR(x;\pprofile,W) \\
x^{*}_{\pprofile,W} \in A^*_{\pprofile, W} & = \argmin_{x \in A} \MR(x;\pprofile,W) 
\end{align}

%Finally,  $\MMR(\pprofile,W)$ is the value of minimax regret under $\pprofile$ and $W$, obtained when recommending a minimax optimal alternative $x^*_{\pprofile, W}$:
%\begin{align*}
%\MMR(\pprofile,W) & = \min_{x \in A} \MR(x;\pprofile,W) \\
%x^{*}_{\pprofile,W} \in A^*_{\pprofile, W} & = \argmin_{x \in A} \MR(x;\pprofile,W) 
%\end{align*}
By picking as consensus choice an alternative associated with minimax regret, we can provide a recommendation that gives worst-case guarantees, giving some robustness in face of uncertainty. 
%In cases of ties in minimax regret, we can either decide to return all minimax alternatives $A^*_{\pprofile, W}$ as winners or to pick just one of them using some tie-breaking strategy.
%
%Notice that if $\MMR(\pprofile, W)\!=\!0$, then $x^{*}_{\pprofile,W}$ is a necessary co-winner; this means that for any valid completion of the profile and any feasible $w \!\in\! W$, $x^{*}_{\pprofile,W}$ obtains a highest score.
%
%We note that our notion of regret gives some cardinal meaning to the scores: instead of just being used to select winners under the corresponding PSR, their differences are considered as representing the regret of the chair.

\section{Interactive Elicitation} 
\label{sec:elicit}
We propose an incremental elicitation method based on minimax regret.
At each step, the system may ask a question either to one of the agents or to the chair.
The goal is to acquire relevant information to reduce minimax regret as quickly as possible.
As termination condition of elicitation, we can check whether minimax regret is lower than a threshold or, if we wish optimality, we can perform elicitation until minimax regret drops to zero.

\paragraph{Question types}
We distinguish between questions asked to the agents and questions asked to the chair.
For the first we consider comparison queries that ask a particular agent to compare two alternatives. Then, the partial profile $\pprofile$ is augmented on the bases of the agent's answer.
The latter, instead, aim at refining our knowledge about the scoring rule; in particular, we want to acquire constraints of the type:
\[ w_{r} - w_{r+1} \geq \lambda (w_{r+1} - w_{r+2}) \] 
for $r \in \{1,\ldots,m-2\}$, relating the difference between the importance of consecutive ranks.
% $r$ and $r+1$ with the difference between ranks $r+1$ and $r+2$.

\paragraph{Elicitation strategies}
A strategy is a function (or a random process) that, given our partial knowledge so far, gives us a question that should be asked.

The \strat{Random} strategy first decides with a probability of $1/2$ each whether it will ask a question about weights or about a preference ordering. Then it equiprobably draws a question among the set of the possible ones. This strategy is clearly non efficient but it gives us a baseline for comparisons.

The \strat{Extreme completions} strategy considers the solution of the minimax regret game and estimates the contribution to the regret of our uncertainty about the weights and the profile. Then it asks a question to the chair or to the agents depending on which of these two values is the highest.

The \strat{Pessimistic} strategy selects among all the possible questions the one that leads to minimal regret in the worst case. For each question, it considers both possible answers to the question and uses some aggregation of their MMRs as a score of the question.

The \strat{Two phase} strategy asks a predefined, non adaptive sequence of $m-2$ questions to the chair and then it only asks questions about the voters, using questions as defined in the \strat{Pessimistic} strategy.

\section{Conclusions}  
\label{sec:conclusions}
In this paper we have considered a social choice setting with partial information and we have proposed the use of minimax regret both as a means of robust winner determination as well as a guide to the process of simultaneous elicitation of preferences and voting rule.

The next step is to test our strategies in order to compare the performance of our interleaved elicitation approach that mix questions to the chair and to the agents, to a more classical approach that elicit the rule first and then the agents preferences (or vice-versa).

Further development of elicitation strategies, considering alternative heuristics, is also an important direction for future works together with the extension of the approach to voting rules beyond scoring rules.

% Acknowledgements: We thank the reviewers for comments helping to improve the paper. 
%{\small
\bibliography{biblio}
\bibliographystyle{plain}
%\bibliographystyle{plain} 
%}
\end{document}
